An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-13 EconGrowth.Rnw
Beginning to add the CES production function. Can't get a solution yet.

<<HIstorical_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.pos="H", fig.cap="Historical data. Indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).">>=
createHistoricalLatticeGraph()
@

@@

<<SFk_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $k$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("K")
@

<<SFl_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $l$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("L")
@

<<SFq_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("Q")
@

<<SFx_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("X")
@

<<SFu_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("U")
@

@@

<<CES functions, eval=FALSE>>=
cesData <- function(countryName, energyType){
  energyColumnName <- paste("i", energyType, sep="")
  # Load the data that we need.
  dataTable <- loadData(countryName)

>>>>

<<Historical_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.pos="H", fig.cap="Historical data. Indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).">>=
createHistoricalLatticeGraph()
@

@@

<<SFk_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $k$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph("K")
@

<<SFl_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $l$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph("L")
@

<<SFq_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph("Q")
@

<<SFx_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph("X")
@

<<SFu_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createSFParamsGraph("U")
@

@@

<<CES functions, eval=TRUE>>=
####################
# Returns an nls ces model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
cesModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadData(countryAbbrev)
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
print(dataTable)
  # Establish guess values for phi beta, zeta, lambda_L and lambda_E.
  phiGuess <- -20
  betaGuess <- 0.5 # a typical value for beta (exponent on labor)
  zetaGuess <- 0.0004 # a small value
  lambda_LGuess <- 0.007 #assuming no technical progress on the labor-capital portion of the function
  lambda_EGuess <- 0.008 #assuming no technical progress on the energy portion of the function
  # Runs a non-linear least squares fit to the data with constraints
  modelCES <- nls(iGDP ~ ((1-zeta) * (exp(lambda_L*iYear) * iCapStk^(1-beta) * iLabor^beta)^phi 
                           + zeta*(exp(lambda_E*iYear) * iEToFit)^phi)^(1/phi), 
                  data=dataTable,
                  start=list(phi=phiGuess, beta=betaGuess, zeta=zetaGuess, lambda_L=lambda_LGuess, 
                             lambda_E=lambda_EGuess),
                  control=nls.control(maxiter=500, tol=1e-06, minFactor=1/1024, 
                                      printEval=FALSE, warnOnly=TRUE),
#                   algorithm = "port",
#                   lower = list(phi=-Inf, beta=0, zeta=0, lambda_L=-Inf, lambda_E=-Inf),
#                   upper = list(phi=0, beta=1, zeta=1, lambda_L=Inf, lambda_E=Inf),
                   )
  return(modelCES)
}
model <- cesModel("UK", "U")
summary <- summary(model)
print(summary)

cesData <- function(countryName, energyType){
  energyColumnName <- paste("i", energyType, sep="")
  # Load the data that we need.
  dataTable <- loadData(countryName)
